### 2012-04-09 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: Corrections.

Our statements are of the three following forms. The first one is $p \rightarrow t$, where $p$ is a PathType representing an XQuery path selecting a set of elements or attributes, and $t$ is an XSD built-in atomic type.

The second and third forms are $k = (C,P,\{L\})$ and $fk = (C(P_1,\{L_1\}) \rightarrow (P_2,\{L_2\}))$, representing a key and a foreign key, where $C,P,P_1,P_2,L,L_1,L_2$ are PathTypes without predicates using only child and descendant axes.

In all three cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($p$ from the first form and $C$ from the second and third form). This is done in two steps. The first one is normalization of a particular PathType and the second one is selection of the targeted elements.

@@

In case of inferred types, the situation is slightly less simple, because there can occur conflicting statements. The examples of such conflicts for path $p$ without predicates are $p_1 \rightarrow date$, $p_2 \rightarrow string$, and, $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.

Note that, in both examples, one type is castable to the other (date to string, and byte to int). Both types are inferred correctly for nodes targeted by $p$, but one of them was inferred from a more convenient expression and is more precise.

Consider these two expressions; PathType $p$ is compared to an integer literal constant, and, PathType $p$ is an argument of function where a formal type of the argument is byte. The first expression is utilized to infer statement $p \rightarrow integer$, and the second one to infer statement $p \rightarrow byte$. Both of the types are correct, but the second one is more accurate.

A problem emerges for example if PathType $p$ is compared to an integer constant, and the real type of elements (or attributes) selected by $p$ is \texttt{double}. In that case, statement $p \rightarrow$ \texttt{integer} is inferred, but it is not correct.

\subsection{Verification using XML data}
To solves those problems, we propose a simple verification using XML data.

For each normalized PathType $p$ from the inferred type statements $\mathcal{S}_t$, we find set $T_p$ of all inferred types. $T_p = \{t|(p \rightarrow t) \in \mathcal{S}_t\}$. Then, we create sequence $T_p'$ by ordering the set $T_p$ from the most specific type to the most general one. For example, if $T_p = \{$\texttt{double}, \texttt{byte}, \texttt{int}$\}$, $T_p' = ($\texttt{byte}, \texttt{int}, \texttt{double}$)$.

Since we have the XML data and path $p$ is an XQuery path, we can use an XQuery processor (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $p$. The verification algorithm iterates through $T_p'$ and for each $t' \in T_p'$ it checks if every node in $N$ conforms to $t'$. If so, $t'$ is the inferred type for nodes $N$ (and PathType $p$), else the inferred type is \texttt{string}.

>>>>

Our statements are of the three following forms. The first one is $P \rightarrow T$, where $P$ is a PathType representing an XQuery path selecting a set of elements or attributes, and $T$ is an XSD built-in atomic type.

The second and third forms are $K = (C,P,\{L\})$ and $K_f = (C(P_1,\{L_1\}) \rightarrow (P_2,\{L_2\}))$, representing a key and a foreign key.

In all three cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($P$ from the first form and $C$ from the second and third form). This is done in two steps. The first one is normalization of a particular PathType and the second one is selection of the targeted elements.

@@

In case of inferred types, the situation is slightly less simple, because there can occur conflicting statements. The examples of such conflicts for path $P$ without predicates are $P_1 \rightarrow$ \texttt{date}, $P_2 \rightarrow$ \texttt{string}, and, $P_1 \rightarrow$ \texttt{byte}, $P_2 \rightarrow$ \texttt{int}, where $P_1$, $P_2$ are paths that when stripped of predicates, they equal $P$.

Note that, in both examples, one type is castable to the other (\texttt{date} to \texttt{string}, and \texttt{byte} to \texttt{int}). Both types are inferred correctly for nodes targeted by $P$, but one of them was inferred from a more convenient expression and is more precise.

Consider these two expressions; PathType $P$ is compared to an integral literal constant, and, PathType $P$ is an argument of a function where a formal type of the argument is \texttt{byte}. The first expression is utilized to infer statement $P \rightarrow$ \texttt{integer}, and the second one to infer statement $P \rightarrow$ \texttt{byte}. Both of the types are correct, but the second one is more accurate.

A problem emerges for example if PathType $P$ is compared to an integral constant, and the real type of elements (or attributes) selected by $P$ is \texttt{double}. In that case, statement $P \rightarrow$ \texttt{integer} is inferred, but it is not correct.

\subsection{Verification using XML data}
To solve the problems, we propose a simple verification using XML data.

For each normalized PathType $P$ from the inferred type statements $\mathcal{S}_t$, we find set $\mathcal{T}_P$ of all inferred types. $\mathcal{T}_P = \{T|(P \rightarrow T) \in \mathcal{S}_t\}$. Then, we create sequence $\mathcal{T}_P'$ by ordering the set $\mathcal{T}_P$ from the most specific type to the most general one. For example, if $\mathcal{T}_P = \{$\texttt{double}, \texttt{byte}, \texttt{int}$\}$, $\mathcal{T}_P' = ($\texttt{byte}, \texttt{int}, \texttt{double}$)$.

Since we have the XML data and path $P$ is an XQuery path, we can use an XQuery processor (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $P$. The verification algorithm iterates through $\mathcal{T}_P'$ and for each $T' \in \mathcal{T}_P'$ it checks if every node in $N$ conforms to $T'$. If so, $T'$ is the inferred type for nodes $N$ (and PathType $P$), else the inferred type is \texttt{string}.
